\begin{abstract}
The FAIR Guiding Principles --- Findable, Accessible, Interoperable,
Reusable --- were drafted for data and have since been extended to
research software (FAIR4RS) and machine-learning models (FAIR4ML).
None of those extensions yet covers the artefact classes that
Large-Language-Model-assisted scientific work has produced: exportable
conversation transcripts, versioned prompt files, tool-and-model version
manifests, verification-status ladders, structured redaction policies,
and per-claim provenance maps. We propose \emph{F(AI)\textsuperscript{2}R},
read \emph{F-A-I-A-I-R} --- the canonical FAIR axes with an additional
\emph{(AI)} factor multiplied through them, squared because each axis
demands two passes over every artefact: an authoring pass and an audit
pass. F(AI)\textsuperscript{2}R is not a new principle. It is the
integration of eight individually unoriginal practices into a single
discipline: transcript preservation, verification-status labelling,
per-claim provenance maps, mirror discipline between markdown and \LaTeX{},
recursive meta-process documentation, base-rate-anchored AI disclosure,
legal honesty about authorship, and FAIR alignment treated as a
precondition rather than a retrofit. None of the eight is novel on its
own; the integration is the contribution. We describe a ten-stage agent
pipeline that operationalises these practices, with scrutinizers
restricted to filing defect registries (the \emph{handback discipline})
so that prose owners and audit owners are never the same agent. We
release this paper, its prompts, and its PROV-O graph as the worked
example.
\end{abstract}
