\section{Prompt Taxonomy}
\label{app:prompts}

Every prompt under \texttt{agents/} follows the same five-section
structure so that a different harness can ingest it with minimal
adaptation.

\begin{description}
    \item[Role.] One paragraph: what the agent is, in plain language,
        with a one-sentence non-goal. The non-goal exists to refuse
        scope creep at read time.
    \item[Primary-artifact consistency (binding).] The shared
        invariant of \S\ref{sec:mirror}, restated with a tailored
        clause for the agent's role: which of (manuscript, graph,
        logbook) the agent updates directly, which it submits
        for update by another agent, and what it must not finish
        without.
    \item[You do.] The actions the agent is permitted to perform, in
        bullet form. Each bullet is small enough to be checked at
        review time.
    \item[You do not.] The actions explicitly out of scope, in bullet
        form. The list is short and load-bearing: an agent that needs
        to be told ``do not invent citations'' is one prompt change
        away from doing so.
    \item[Output.] The artefact the agent produces and where it lands.
        Outputs are typed: a draft is a diff, an audit is a defect
        registry, a graph update is a triple bundle, a section file is
        \texttt{paper/sections/<slug>.tex}.
\end{description}

\paragraph{Granularity rules.}
\begin{itemize}
    \item One agent per role; one role per file.
    \item One section per \texttt{.tex} file under
        \texttt{paper/sections/}; one \texttt{.tex} per section. Soft
        target: 100--400 lines per file. The
        \texttt{scientific-writer} prompt enforces this.
    \item One activity per stage in \texttt{doc/provenance.ttl};
        an activity that runs more than once gets repeated entries
        with distinct \texttt{prov:startedAtTime}.
\end{itemize}

\paragraph{Refusal conditions.}
Each prompt names at least one refusal condition. Common ones:
\emph{do not invent citations}; \emph{do not write to the graph if a
submission lacks a corresponding manuscript change}; \emph{do not
deploy the public site without recorded publication consent}. Refusal
is a feature, not a defect.
