\section{Process evolution: the chronology}
\label{sec:appendix-evolution}

This appendix expands the digest in
Table~\ref{tab:evolutions} (\S\ref{sec:evolution}) into a per-step
chronology. The body of the paper carries the table because the
table is what a reader needs at submission time; the chronology
below is the longer record, kept here so the table is auditable
against narrative without expanding the body.


\paragraph{Bootstrap.}
Researcher request: initialise a paper-plus-pipeline repository with
prompts under \texttt{agents/}, a logbook under \texttt{doc/}, and a
single PROV-O graph that follows the creation of the paper, distilled
from a prior repository (\texttt{noheton/Obscurity-Is-Dead}) with all
domain content abstracted away. The first AI draft read
F(AI)\textsuperscript{2}R as ``AI participates twice in the loop''.
Reading the prior repository carefully revealed the researcher's
original framing: \emph{(AI)} as a factor multiplied through every
FAIR axis, squared because of the two passes per axis. The paper
draft adopts the latter. Lesson: when the AI's first-pass
interpretation differs from the researcher's prior writing, the prior
writing wins.

\paragraph{Primary-artefact consistency invariant.}
After the bootstrap, the researcher asserted that the manuscript,
the PROV-O graph, and the logbook are primary artefacts and must
remain consistent and up to date at all times. The AI added a
binding ``Primary-artefact consistency'' block to every prompt;
named the Orchestrator the ultimate custodian and the Provenance
Curator the keystone; specified that the FAIR Aligner surfaces a
desync as an audit \emph{fail}. The rule was recorded as a
\texttt{fair2r:Claim}, attributed to the researcher, and as a
\texttt{prov:Activity} revising every affected prompt entity.

\paragraph{Static site, Pages workflow, paper-build CI.}
Researcher request: a public site in the spirit of the prior
repository, plus continuous integration. The AI added a static-site
generator (parsing \texttt{doc/provenance.ttl} via \texttt{rdflib}),
the corresponding GitHub Pages workflow with a Turtle-validation
gate, and a paper-build workflow.

\paragraph{Chapter granularity for the writer.}
Researcher request: \emph{each \texttt{.tex} file should represent
one chapter}. The Scientific Writer prompt gained a binding
``File granularity'' block: one \texttt{\textbackslash section\{\}}
per file, soft target 100--400 lines, filename slug matches the
\texttt{\textbackslash label\{sec:\dots\}}. \texttt{paper/main.tex}
became a thin assembler.

\paragraph{Recursive attestation, by request.}
Researcher request: \emph{add a section that this paper is generated
using exactly this process}. \S\ref{sec:self} was added.

\paragraph{This section, by request.}
Researcher request: \emph{document the evolution of the process
during cooperative writing}. This section. The \texttt{fair2r:Claim}
entries for it sit at \textsc{ai-confirmed} until reviewed.

\paragraph{Short-form pivot.}
Researcher request: write only the short form (10 pages of body plus
references and appendices) rather than maintaining a long/condensed
split. The \texttt{main-condensed.tex} target was retired; the
Condenser agent was repurposed as a \emph{page-budget enforcer} that
fires when the compiled PDF exceeds the configured budget; the
Scientific Writer prompt gained a page-budget rule.

\paragraph{Institutional design system.}
Researcher request: integrate the DLR design system (Frutiger /
Arial, blue/green/yellow chapter accents, square corners, hairline
borders, photography-led, no marketing register). The site CSS was
rewritten on top of the DLR \texttt{colors\_and\_type.css} tokens; a
small \LaTeX{} style file (\texttt{paper/style/fair2r.sty}) carries
the equivalent typography and rule colours into the PDF; the
imprint was updated with the author's institutional identity (DLR
ZLP Augsburg, Helmholtz, NFDI4Ing, HMC).

\subsection*{Patterns observed}

Three patterns recur and are worth naming.

\paragraph{Researcher-as-corrector.}
The researcher's interventions are short, specific, and rule-shaped.
None is a rewrite of AI-produced prose. The cooperation lives in the
gap between a model that produces a plausible default and a
researcher who can recognise when the default is wrong.

\paragraph{Pipeline as target, not yet as practice.}
The ten-stage pipeline is the architecture the prompts encode. The
present draft was produced by a single AI agent in conversation with
the researcher, with role separation documented in the prompts but
not yet actively switched per turn. We name this gap rather than
paper over it.

\paragraph{Logbook as conscience.}
Every event above corresponds to an entry in
\texttt{doc/logbook.md}. The append-only discipline made the
reconstruction cheap. \emph{The logbook is the conscience of the
cooperation; the provenance graph is its record.}

\subsection*{Field notes from the second cooperative writing project}

The methodological ancestor~\cite{krebs2026obscurity} was the
author's first cooperative-writing project. The present paper is the
second. Three observations from the second-time experience are worth
recording; the full set lives in
\texttt{doc/user-observations-log.md}.

\paragraph{The researcher's role is shifting toward experimentation.}
Cooperative writing with an LLM, as practised here, lifts the
prose-production load off the researcher and leaves more cycles for
abstraction, decision-making, and design of the next experiment. For
empirical disciplines this is a net good: writing-as-administrative-
overhead shrinks; experimentation-as-the-actual-contribution grows.
For paper-driven disciplines --- philosophy, parts of the humanities
--- the picture is more uncomfortable: the prose \emph{is} the
experiment, and removing the slow, deliberate craft of writing risks
removing the thinking that the writing was
doing~\cite{kuteeva2024diversity}\todo[inline]{verify}. We name the
asymmetry rather than pretend F(AI)\textsuperscript{2}R is
discipline-neutral.

\paragraph{Prompt-craft is closer to programming than to drafting.}
A second-time observation: writing a prompt feels easier than writing
the corresponding paragraph oneself. The short conversational cycle
maps, for a researcher trained in computer science, to the familiar
loop of writing a program and watching its output --- closer to
experimentation than to formal academic
writing~\cite{reynolds2021prompt,liu2023prompt}\todo[inline]{verify}.
This is a feature (it lowers the activation energy of getting
started, which is itself a contribution to throughput) and a risk
(prompt-craft can become a comfortable substitute for the harder
parts of writing). The risk is managed, in
F(AI)\textsuperscript{2}R, by keeping the human in the deciding seat
(\S\ref{sec:authors-note}, \emph{What surprised me about the
cooperation}) and by the per-claim provenance map that makes drift
detectable.

\paragraph{Responsibility includes the rooms after publication.}
The model-side contribution ends at the commit. The human-side
contribution continues into the rooms where the work has to be
defended in person --- conferences, panels, discussions, reviewer
correspondence. AI cannot share that load: the model cannot consent,
cannot carry a reputation, and cannot be cross-examined at the next
RDA plenary or HMC symposium. The seventh integrated practice
(\S\ref{sec:eight}, item~7) names the legal version of this point;
the field-notes version is more ordinary. \emph{Whoever puts their
name on the paper signs up for every conversation the paper produces
afterwards}, and the AI's contribution is, by construction, absent
from those conversations. The interpersonal-sharing leg of authorship
is the leg the methodology section cannot eliminate even in
principle.
