\section{Provenance analysis: queries and worked numbers}
\label{sec:appendix-provenance-analysis}

The body section (\S\ref{sec:provenance-analysis}) summarises the
audit; this appendix walks the four diagnostic queries on the live
graph. The auto-generated tables below are produced by
\texttt{scripts/provenance\_analysis.py}, which the build pipeline
re-runs on every commit. They reflect the state of
\texttt{doc/provenance.ttl} at the time this PDF was compiled and
are embedded into the prose by the build, not transcribed by hand.

\paragraph{Verification rung distribution.}
Every claim in the graph carries a
\texttt{fair2r:verificationState}; we report the fraction at each
rung at submission time per the sixth integrated practice
(\S\ref{sec:eight}, item~6). The query is one SPARQL pattern:

\begin{quote}\small
\texttt{SELECT ?state (COUNT(?c) AS ?n)\\
WHERE \{ ?c a fair2r:Claim ; fair2r:verificationState ?state \}\\
GROUP BY ?state}
\end{quote}

\noindent The result for the present manuscript:

\input{sections/_generated/rung-distribution.tex}

\paragraph{Section-to-claim coverage.}
Every section in \texttt{paper/main.tex} should appear in the graph
as a \texttt{fair2r:Section} entity, with at least one
\texttt{prov:Activity} that produced it and (almost always) at least
one \texttt{fair2r:Claim} attributed to that activity. A section
that exists in prose but not in the graph is a defect surfaced by
the FAIR Aligner.

\input{sections/_generated/section-coverage.tex}

\noindent A coverage row that reads \emph{prose: yes / graph: no} is
a hard fail; \emph{graph: yes / claims: 0} is a soft warning meaning
the section has been registered but the claim curation has not
caught up.

\paragraph{Source invocation strength.}
Every \texttt{fair2r:Source} entity that the manuscript cites should
appear as the object of at least one
\texttt{prov:wasDerivedFrom} edge from a \texttt{fair2r:Claim}. A
source cited but never linked from a claim is most often a
\textbf{ghost citation} --- a reference that lives in the
bibliography because the prose mentions it, but that no claim
actually relies on. In a healthy F(AI)\textsuperscript{2}R
manuscript, the count is zero.

\input{sections/_generated/source-strength.tex}

\paragraph{Activity timeline.}
The \texttt{prov:Activity} nodes form a dated record of every
substantive change. The most recent five activities, in
ISO-8601 order, are:

\input{sections/_generated/recent-activities.tex}

\noindent The same data, with one row per session, is the
human-readable mirror in \texttt{doc/logbook.md}; drift between the
two is itself a defect (see \S\ref{sec:handback} on the handback
discipline).

\paragraph{Keep this appendix honest.}
This appendix is generated against the current graph. If a future
revision adds a verification rung, a new section, or a new failure
mode, the corresponding entry here updates automatically on the next
build; the prose around the figures stays put. The intent is that a
reader who clones the repository, recompiles the PDF, and re-reads
this appendix, will see the analysis reflect the manuscript that was
actually produced --- not a snapshot frozen at draft time.
