\section{Background}
\label{sec:background}

We assume the reader is broadly familiar with FAIR. This section fixes
notation, reviews the two prior FAIR extensions on which we lean, and
introduces the three further frameworks --- PROV-O, PRISMA 2020, and
GRADE --- whose combination, we will argue, supplies the missing
machinery for evidence-anchored LLM-assisted scholarship.

\subsection{FAIR}
\label{sec:bg-fair}

The FAIR Guiding Principles --- Findable, Accessible, Interoperable,
Reusable --- were published by Wilkinson and
colleagues~\cite{wilkinson2016fair}. The principles are organised into
fifteen sub-principles (F1--F4, A1--A2 with A1.1, A1.2 and A2, I1--I3,
R1 with R1.1--R1.3). FAIR is, by design, a guideline set rather than a
specification: it describes the properties that make data fit for
machine consumption without prescribing the technologies that supply
those properties. The deliberate looseness is what has let the
principles travel.

The looseness has also made FAIR easy to retrofit poorly. A common
failure pattern is to declare FAIR compliance after the fact, when
identifiers have been minted by accident, machine-readable metadata is
generated from a brittle scrape of the manuscript, and the
``interoperable'' and ``reusable'' sub-principles are checked by
intention rather than evidence. Section~\ref{sec:eight} returns to this
as the eighth integrated practice: \emph{do not retrofit FAIR}.

\subsection{FAIR4RS and FAIR4ML}
\label{sec:bg-fair-extensions}

FAIR4RS~\cite{chuehong2022fair4rs} re-reads each of the FAIR
sub-principles for research software. The exercise is non-trivial:
software has versions and dependencies that data does not have in the
same way, and the ``reusable'' axis carries license, provenance, and
build-reproducibility weight that the original FAIR formulation
under-specifies. FAIR4ML~\cite{ravi2024fair4ml} performs the same
exercise for machine-learning models, where ``reusable'' picks up
training-data documentation, evaluation regimes, and intended-use
statements derived from the model-cards and datasheets
literature~\cite{mitchell2019modelcards,gebru2021datasheets}.

Two observations:

\begin{enumerate}
    \item Each extension preserves the four FAIR letters and adds new
        sub-principles or refines existing ones for the new artefact
        class. The pattern is additive, not revisionist.
    \item Neither extension explicitly accommodates artefact classes
        that exist downstream of \emph{the writing process itself} ---
        prompts, transcripts, verification ladders, redaction policies,
        and per-claim provenance maps.
\end{enumerate}

F(AI)\textsuperscript{2}R is the third such extension. It follows
FAIR4RS and FAIR4ML in being additive. It differs in that the
artefacts it covers are processual rather than terminal: a prompt is
not the output of the research, but a versioned input that determines
how the output came about.

\subsection{PROV-O}
\label{sec:bg-prov}

The W3C PROV family~\cite{w3c2013provo} supplies a vocabulary for
expressing provenance: \texttt{prov:Entity} (something that exists),
\texttt{prov:Activity} (something that happens, possibly to entities),
\texttt{prov:Agent} (something or someone responsible). The relations
that matter most for our purposes are \texttt{prov:wasGeneratedBy},
\texttt{prov:wasAttributedTo}, \texttt{prov:wasDerivedFrom}, and
\texttt{prov:hadPlan}. PROV-O is the OWL profile and is sufficient for
the level of detail we need.

We introduce a small project namespace
(\texttt{fair2r:}) with subclasses for the artefacts particular to
LLM-assisted scholarship --- \texttt{fair2r:Claim},
\texttt{fair2r:Source}, \texttt{fair2r:Prompt} (a
\texttt{prov:Plan}), \texttt{fair2r:AIAgent} and
\texttt{fair2r:HumanResearcher} (subclasses of
\texttt{prov:SoftwareAgent} and \texttt{prov:Person} respectively),
\texttt{fair2r:AuthoringPass} and \texttt{fair2r:AuditPass}
(subclasses of \texttt{prov:Activity}), and a
\texttt{fair2r:VerificationState} class whose individuals are the
rungs of the ladder defined in Section~\ref{sec:ladder}. The full
schema is in \texttt{doc/provenance.ttl} of the accompanying
repository; Appendix~\ref{app:fair-cells} shows it as a table.

PROV-O is necessary but not sufficient. It records that an activity
happened and that an agent participated; it does not, on its own, say
how confidently a claim has been verified. That is what the
verification ladder is for.

\subsection{PRISMA 2020 and GRADE}
\label{sec:bg-prisma-grade}

PRISMA 2020~\cite{page2021prisma} is the dominant reporting standard
for systematic reviews. Its flow diagram (Identification, Screening,
Included) is, structurally, a verification ladder: each rung represents
a deeper engagement with the literature, and an item's rung at
publication time tells the reader how much retrieval effort underpins
its inclusion.

GRADE~\cite{guyatt2008grade} is the dominant framework for rating the
\emph{certainty of evidence} behind a clinical recommendation. It is
not a retrieval ladder but a strength-of-evidence ladder:
\textsc{very low}, \textsc{low}, \textsc{moderate}, \textsc{high}.

We will argue, in Section~\ref{sec:ladder}, that the verification
ladder F(AI)\textsuperscript{2}R uses for individual claims is best
read as a crosswalk between PRISMA 2020 (retrieval depth) and GRADE
(invocation strength). The crosswalk is the floor a rung permits, not
a guarantee of the ceiling.

\subsection{AI-disclosure norms}
\label{sec:bg-norms}

The disclosure landscape is a moving target. The \textsc{ICMJE}
Recommendations and most major venues now require authors to disclose
the use of generative AI tools and to take responsibility for the
output~\cite{icmje2023}. Some venues have explicit policies on
AI co-authorship, all of which converge on one position: AI is not an
author. The legal background, in jurisdictions including
Germany (\S2 \emph{UrhG}), is consistent: AI cannot hold copyright,
cannot consent to publication, and cannot be held accountable. We
return to this in the seventh integrated practice
(\S\ref{sec:practice-legal}).

The norms cover \emph{whether to disclose}. They do not cover
\emph{what to preserve}: which transcripts, which prompt versions,
which model identifiers, which verification states. F(AI)\textsuperscript{2}R
is the proposal that there is a usable answer.
