\section{Conclusion}
\label{sec:conclusion}

The FAIR Guiding Principles were drafted for data and have aged
well. The artefacts that LLM-assisted scholarly writing produces are
not data, not software, and not models; they need their own
re-reading. F(AI)\textsuperscript{2}R is that re-reading: the
canonical FAIR axes with an \emph{(AI)} factor multiplied through
them, squared because each axis demands two passes over every
artefact.

The contribution is not the name. The contribution is the
integration of eight individually unoriginal practices ---
transcript preservation, verification-status labelling, per-claim
provenance maps, mirror discipline, recursive meta-process
documentation, base-rate-anchored disclosure, legal honesty about
authorship, and FAIR alignment as precondition --- enforced by a
ten-stage agent pipeline whose prose-owning and audit-owning roles
are strictly separated through a handback discipline. The
integration is the contribution.

\paragraph{Submission conditions.}
We will submit F(AI)\textsuperscript{2}R to the Research Data
Alliance and to FORCE11 as a candidate community recommendation
when two preconditions are met: (i) human-author consent for
community-recommendation submission is recorded in
\texttt{doc/publication-consent.md}; and (ii) at least one external
case study has adopted the pattern on a paper unrelated to either
the source repository or this one, and the resulting artefacts are
public under a permissive licence. The conditions are the minimum
we need to distinguish a single-author idiosyncrasy from a
transferable pattern.

\paragraph{Call to action.}
We release this paper, its prompts, its provenance graph, its build
pipeline, and its public site as the worked example, and we invite
a second case study --- by a different author, on a different
paper, in a different domain --- as the precondition for
community-recommendation submission. The repository is open and
the licences are permissive. A successor framework that argues
with this one, replaces it, or tightens it is not a refutation; it
is the point.

\paragraph{The longer arc.}
F(AI)\textsuperscript{2}R is not the destination, only the entry
ticket. The bioinformatics precedent
(\companionref{sec:appendix-pattern-extensions}{Appendix~G:
Pattern Extensions}) suggests where the
trajectory leads: under the volume pressure that
LLM-assisted scholarship is now imposing, future research success
will depend less on the solitary craft of a single manuscript and
more on three properties this schema is designed to support ---
provenance models that turn published claims into machine-readable
hypotheses for the next study, graph interconnectivity across
institutional knowledge graphs (a fully-instrumented Helmholtz
cross-centre graph being the proximate
example~\cite{curdt2025hmc}), and clever experimental validation
that uses the graph to identify the next testable deviation from
the published consensus. The eight integrated practices serve all
three. We are not arguing F(AI)\textsuperscript{2}R has to be the
standard; we are arguing that some standard with these properties
is now load-bearing for the next decade of scholarship, and that
naming the practice we are already drifting toward is the cheapest
way to make it contestable in the open literature.

\paragraph{The bridge to what is now unthinkable.}
Xennials --- born late enough to grow up without the internet and
early enough to spend their professional life inside
it~\cite{shafer2014xennials}\todo[inline]{verify} --- remember a
working day whose discipline came not from nostalgia but from
constraint: each step cost something, and the costs shaped the
questions. Senior researchers carry an analogous responsibility:
to articulate, in machine-readable form, what was previously
unthinkable to question, so that the practice that was
load-bearing in our training survives translation into a toolset
that no longer demands it. The eight integrated practices are
the first crossing; the ladder, the PROV-O graph, and the
handback discipline are its rails. The next cohort, raised inside
a tooling whose default is fluency, deserves to know what made
the discipline necessary before the choice to keep it or replace
it is theirs.

\bigskip
\noindent \emph{The pipeline is the team; the team is the
methodology. Specialisation reduces hallucination at the seam.
Naming what we are already doing is the first step toward
surrendering it to peer scrutiny.}
