\section{Discussion}
\label{sec:discussion}

\subsection{What F(AI)\texorpdfstring{\textsuperscript{2}}{2}R is not}

It is not a vendor stack. The pattern is indifferent to which model is
behind a given agent, which harness invokes it, or which serialisation
format the provenance graph takes. Substituting Turtle for JSON-LD,
GitHub Actions for any other CI, or any one model for any other,
leaves the eight practices intact.

It is not a guarantee. A team that adopts the eight practices and
the pipeline can still ship a bad paper. The pattern reduces specific
failure modes to specific remediations and surfaces residual ones; it
does not promise that what survives is true.

It is not the first attempt at any of the eight. Datasheets for
datasets~\cite{gebru2021datasheets}, model
cards~\cite{mitchell2019modelcards}, FAIR4RS~\cite{chuehong2022fair4rs},
and the broader research-software-engineering community have spent the
last decade publishing the precursors. The integration is what we
claim is new; we claim it modestly.

\subsection{Limitations of a single worked example}

The pattern has been exercised on two papers (the source repository
and this one), both authored by the same hobbyist researcher. The
sample is too small to support claims about the pattern's behaviour at
scale, in adversarial review settings, or in collaborations whose
contributors do not share the same toolchain. We expect the
verification-overhead figures to be sensitive to the heaviness of the
audit pass and the maturity of the harness; we do not have data to
quantify the dependence.

A community case study --- two or more independent groups, each
applying the pattern to a different paper --- is the prerequisite for
a candidate community recommendation. Until that exists, the
contribution is a worked example with explicit invariants, offered for
adoption rather than for adoption-by-default.

\subsection{Submission conditions}

We will submit F(AI)\textsuperscript{2}R to the
Research Data Alliance and to FORCE11 as a candidate community
recommendation when (and only when) two preconditions are met:
\begin{enumerate}
    \item Human-author consent for community-recommendation submission
        is recorded as a signed entry in
        \texttt{doc/publication-consent.md}.
    \item At least one external case study has adopted the eight
        practices and the pipeline on a paper that is unrelated to
        either the source repository or this one, and the resulting
        artefacts have been made public under a permissive licence.
\end{enumerate}
The conditions are not bureaucratic. They are the minimum we need to
distinguish a single-author idiosyncrasy from a transferable pattern.

\subsection{The dual-use surface}

We are explicit about the dual-use surface (\S\ref{sec:fm-dualuse}).
A pipeline of prompts, a finite-state ladder, and a structured
provenance schema are an instruction set that can be adopted by
bad-faith authors as readily as by good-faith ones. F(AI)\textsuperscript{2}R
\emph{makes the dual-use surface visible} in a form that can be
reasoned about; it does not eliminate it. The mitigation, in our
view, is community: a bad-faith adoption is detectable by inspection
of the artefacts the pattern requires, in a way that a bad-faith
opaque process is not.

\subsection{On AI co-authorship, again}

We have rehearsed this in \S\ref{sec:practice-legal}. The position
holds. AI cannot hold copyright, cannot consent, cannot be held
accountable. The pattern accommodates AI as a first-class
\texttt{prov:Agent} in the graph and as an acknowledged contributor in
the prose; it never accommodates AI as an author, whatever the
incentive. The line is not symbolic. \emph{Legal honesty about
authorship is the practice that keeps every other practice usable.}

\subsection{Naming the practice}

Naming what we are already doing is the first step toward surrendering
it to peer scrutiny. The eight practices were not invented for this
paper; they are the practices the source repository's author had
gradually adopted by attrition, after each individually-unoriginal
practice paid for itself against a specific failure mode. The act of
naming, here, does two things at once: it makes the practices
contestable in the open literature, and it lets a third party adopt
them without re-deriving each from the failure modes that motivated
it. Both effects are the contribution.
