\section{Process Evolution During Cooperative Writing}
\label{sec:evolution}

\S\ref{sec:self} attests that this paper was produced by the
F(AI)\textsuperscript{2}R pipeline. This section is the second half
of that attestation: a candid record of how the process itself
changed while the paper was being written. The goal is transparency
about how the paper came to be and how researcher and AI interacted.
The record is reconstructed from \texttt{doc/logbook.md} and
\texttt{doc/provenance.ttl}.

\paragraph{Summary of evolutions.}
Table~\ref{tab:evolutions} synthesises the changes the cooperation
imposed on its own scaffolding while the paper was being written.
Five classes of evolution show up: \emph{rules} (binding constraints
added to one or more agent prompts), \emph{agents} (a new specialised
agent added to \texttt{agents/}), \emph{schema} (a class or property
added to \texttt{doc/provenance.ttl}), \emph{pipeline} (a build,
deployment, or rendering change), and \emph{manuscript} (a structural
reframing of the paper itself). The full chronology lives in
\texttt{doc/logbook.md}; the table below is a deduplicated digest of
the structurally-significant evolutions and is auditable against the
forty-eight typed entries in \texttt{doc/user-contributions.md}.

\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{1.55cm}p{1.4cm}p{4.2cm}p{6.6cm}@{}}
\caption{Process evolutions during cooperative writing. \emph{Class}:
rule (binding constraint), agent (new specialised prompt), schema
(graph extension), pipeline (build / render / deploy change),
manuscript (structural reframing). \emph{Trigger / impact}: the
researcher prompt or audit finding that landed the change and what it
unlocked downstream. Reconstructed from \texttt{doc/logbook.md} and
\texttt{doc/user-contributions.md}.}
\label{tab:evolutions} \\
\toprule
When & Class & What changed & Trigger / impact \\
\midrule
\endfirsthead
\toprule
When & Class & What changed & Trigger / impact \\
\midrule
\endhead
\bottomrule
\endfoot
2026-05-06 & Rule       & Primary-artefact consistency invariant (manuscript, graph, logbook, slides must agree). & Researcher direction; orchestrator named keystone; FAIR-aligner surfaces desync as audit \texttt{fail}. \\
2026-05-06 & Rule       & Chapter-per-file rule for \texttt{.tex}.            & \texttt{paper/main.tex} refactored to a thin assembler; one \verb|\section| per file; enabled chapter-level handback. \\
2026-05-06 & Rule       & Page budget (\(\sim\)10\,pp body) and short-form pivot. & Researcher framing as a position paper; condenser agent repurposed as page-budget enforcer. \\
2026-05-06 & Rule       & Source-research mandate; Consensus / Scholar default. & Cut hallucinated-citation surface; paywall-escalation sub-rule shipped together. \\
2026-05-06 & Rule       & Contribution-tracking rule (this file's existence). & Researcher question \emph{``what is my contribution?''}; spawned \texttt{user-contributions.md} + \texttt{fair2r:Contribution} schema branch. \\
2026-05-06 & Schema     & \texttt{fair2r:Contribution} class with seven sub-types. & Operationalises the partition the Author's Note describes; mirrors per-side audit. \\
2026-05-06 & Schema     & \texttt{ai-confirmed} verification rung (between \texttt{lit-retrieved} and \texttt{lit-read}). & Felt cost of single-tier verification; rung added rather than weakening the existing two. \\
2026-05-06 & Schema     & \texttt{fair2r:Slidedeck} class; slides as fourth primary artefact. & Researcher request for two decks; \texttt{presentation} agent owns them. \\
2026-05-06 & Agent      & \texttt{presentation} agent added to \texttt{agents/}. & Slides became binding artefact; new prompt minted rather than overloading \texttt{scientific-writer}. \\
2026-05-06 & Manuscript & Position-paper reframing (Position + Objections sections). & Researcher direction \emph{``fair should be more a position paper''}; reshaped the contribution narrative. \\
2026-05-06 & Manuscript & Author's Note carried over from \texttt{noheton/Obscurity-Is-Dead}. & Voice-and-stance anchor; the methodology's first-person register lives there only. \\
2026-05-06 & Pipeline   & Static site builder \(+\) Pages workflow \(+\) paper-build CI. & Made the artefact public from day one; cache-bust on the deployed site followed when staleness surfaced. \\
2026-05-06 & Pipeline   & Marp slide pipeline retired; Beamer-only. & Researcher direction \emph{``only ship beamer''}; one toolchain to maintain instead of two. \\
2026-05-07 & Rule       & DLR Corporate Design pinned for every illustrator tool. & Default tool theming counts as an unfinished figure; written above the tool list, not below. \\
2026-05-07 & Rule       & Illustration toolset opened beyond TikZ (matplotlib, mermaid, networkx, graphviz, \dots). & Researcher direction; \texttt{ladder-populations} and \texttt{contribution-histogram} matplotlib figures unlocked. \\
2026-05-07 & Rule       & No-parentless-claim rule (every claim must name a generating activity). & Eight-claim defect found by the verification example; rule landed by the curator pass to prevent recurrence. \\
2026-05-07 & Schema     & \texttt{fair2r:repairs} object property. & Curator pass: per-claim repair activity carries the reason as \texttt{rdfs:comment}; reasons are queryable. \\
2026-05-07 & Agent      & \texttt{research-protocol} agent added to \texttt{agents/}. & Standalone prompt for verification-axis research notes; produced \texttt{provenance-verification.md}. \\
2026-05-07 & Pipeline   & Pages source switched from branch deploy to GitHub Actions. & Default branch became \texttt{main}; CI is single source of truth for the live site. \\
2026-05-07 & Pipeline   & LaTeX-fragment stripper in the site builder. & Markdown sources embedded \verb|\emph{}|, \verb|\textsc{}|, \verb|\begin{enumerate}|; the rendered Pages site needed a clean prose pass. \\
\end{longtable}

\paragraph{Bootstrap.}
Researcher request: initialise a paper-plus-pipeline repository with
prompts under \texttt{agents/}, a logbook under \texttt{doc/}, and a
single PROV-O graph that follows the creation of the paper, distilled
from a prior repository (\texttt{noheton/Obscurity-Is-Dead}) with all
domain content abstracted away. The first AI draft read
F(AI)\textsuperscript{2}R as ``AI participates twice in the loop''.
Reading the prior repository carefully revealed the researcher's
original framing: \emph{(AI)} as a factor multiplied through every
FAIR axis, squared because of the two passes per axis. The paper
draft adopts the latter. Lesson: when the AI's first-pass
interpretation differs from the researcher's prior writing, the prior
writing wins.

\paragraph{Primary-artefact consistency invariant.}
After the bootstrap, the researcher asserted that the manuscript,
the PROV-O graph, and the logbook are primary artefacts and must
remain consistent and up to date at all times. The AI added a
binding ``Primary-artefact consistency'' block to every prompt;
named the Orchestrator the ultimate custodian and the Provenance
Curator the keystone; specified that the FAIR Aligner surfaces a
desync as an audit \emph{fail}. The rule was recorded as a
\texttt{fair2r:Claim}, attributed to the researcher, and as a
\texttt{prov:Activity} revising every affected prompt entity.

\paragraph{Static site, Pages workflow, paper-build CI.}
Researcher request: a public site in the spirit of the prior
repository, plus continuous integration. The AI added a static-site
generator (parsing \texttt{doc/provenance.ttl} via \texttt{rdflib}),
the corresponding GitHub Pages workflow with a Turtle-validation
gate, and a paper-build workflow.

\paragraph{Chapter granularity for the writer.}
Researcher request: \emph{each \texttt{.tex} file should represent
one chapter}. The Scientific Writer prompt gained a binding
``File granularity'' block: one \texttt{\textbackslash section\{\}}
per file, soft target 100--400 lines, filename slug matches the
\texttt{\textbackslash label\{sec:\dots\}}. \texttt{paper/main.tex}
became a thin assembler.

\paragraph{Recursive attestation, by request.}
Researcher request: \emph{add a section that this paper is generated
using exactly this process}. \S\ref{sec:self} was added.

\paragraph{This section, by request.}
Researcher request: \emph{document the evolution of the process
during cooperative writing}. This section. The \texttt{fair2r:Claim}
entries for it sit at \textsc{ai-confirmed} until reviewed.

\paragraph{Short-form pivot.}
Researcher request: write only the short form (10 pages of body plus
references and appendices) rather than maintaining a long/condensed
split. The \texttt{main-condensed.tex} target was retired; the
Condenser agent was repurposed as a \emph{page-budget enforcer} that
fires when the compiled PDF exceeds the configured budget; the
Scientific Writer prompt gained a page-budget rule.

\paragraph{Institutional design system.}
Researcher request: integrate the DLR design system (Frutiger /
Arial, blue/green/yellow chapter accents, square corners, hairline
borders, photography-led, no marketing register). The site CSS was
rewritten on top of the DLR \texttt{colors\_and\_type.css} tokens; a
small \LaTeX{} style file (\texttt{paper/style/fair2r.sty}) carries
the equivalent typography and rule colours into the PDF; the
imprint was updated with the author's institutional identity (DLR
ZLP Augsburg, Helmholtz, NFDI4Ing, HMC).

\subsection*{Patterns observed}

Three patterns recur and are worth naming.

\paragraph{Researcher-as-corrector.}
The researcher's interventions are short, specific, and rule-shaped.
None is a rewrite of AI-produced prose. The cooperation lives in the
gap between a model that produces a plausible default and a
researcher who can recognise when the default is wrong.

\paragraph{Pipeline as target, not yet as practice.}
The ten-stage pipeline is the architecture the prompts encode. The
present draft was produced by a single AI agent in conversation with
the researcher, with role separation documented in the prompts but
not yet actively switched per turn. We name this gap rather than
paper over it.

\paragraph{Logbook as conscience.}
Every event above corresponds to an entry in
\texttt{doc/logbook.md}. The append-only discipline made the
reconstruction cheap. \emph{The logbook is the conscience of the
cooperation; the provenance graph is its record.}

\subsection*{Field notes from the second cooperative writing project}

The methodological ancestor~\cite{krebs2026obscurity} was the
author's first cooperative-writing project. The present paper is the
second. Three observations from the second-time experience are worth
recording; the full set lives in
\texttt{doc/user-observations-log.md}.

\paragraph{The researcher's role is shifting toward experimentation.}
Cooperative writing with an LLM, as practised here, lifts the
prose-production load off the researcher and leaves more cycles for
abstraction, decision-making, and design of the next experiment. For
empirical disciplines this is a net good: writing-as-administrative-
overhead shrinks; experimentation-as-the-actual-contribution grows.
For paper-driven disciplines --- philosophy, parts of the humanities
--- the picture is more uncomfortable: the prose \emph{is} the
experiment, and removing the slow, deliberate craft of writing risks
removing the thinking that the writing was
doing~\cite{kuteeva2024diversity}\todo[inline]{verify}. We name the
asymmetry rather than pretend F(AI)\textsuperscript{2}R is
discipline-neutral.

\paragraph{Prompt-craft is closer to programming than to drafting.}
A second-time observation: writing a prompt feels easier than writing
the corresponding paragraph oneself. The short conversational cycle
maps, for a researcher trained in computer science, to the familiar
loop of writing a program and watching its output --- closer to
experimentation than to formal academic
writing~\cite{reynolds2021prompt,liu2023prompt}\todo[inline]{verify}.
This is a feature (it lowers the activation energy of getting
started, which is itself a contribution to throughput) and a risk
(prompt-craft can become a comfortable substitute for the harder
parts of writing). The risk is managed, in
F(AI)\textsuperscript{2}R, by keeping the human in the deciding seat
(\S\ref{sec:authors-note}, \emph{What surprised me about the
cooperation}) and by the per-claim provenance map that makes drift
detectable.

\paragraph{Responsibility includes the rooms after publication.}
The model-side contribution ends at the commit. The human-side
contribution continues into the rooms where the work has to be
defended in person --- conferences, panels, discussions, reviewer
correspondence. AI cannot share that load: the model cannot consent,
cannot carry a reputation, and cannot be cross-examined at the next
RDA plenary or HMC symposium. The seventh integrated practice
(\S\ref{sec:eight}, item~7) names the legal version of this point;
the field-notes version is more ordinary. \emph{Whoever puts their
name on the paper signs up for every conversation the paper produces
afterwards}, and the AI's contribution is, by construction, absent
from those conversations. The interpersonal-sharing leg of authorship
is the leg the methodology section cannot eliminate even in
principle.
