\section{Failure Modes Named and Addressed}
\label{sec:failure-modes}

The pattern is meant to resist a finite set of failure modes that the
literature on LLM-assisted writing has documented. We name them, say
which practice mitigates each, and are honest about which the pattern
does not address.

\subsection{Fabricated citations}
\label{sec:fm-fabrication}

Reported base rates of fabricated citations in LLM output range from
roughly 18\% in general-purpose biomedical settings to 55\% in
specialised legal queries~\cite{walters2023fabrication,magesh2024legal}.
Fabrication includes: an entirely invented title, an invented DOI for a
real title, a real DOI attached to a wrong title, and a real citation
attached to a claim the source does not in fact make.

The verification ladder (\S\ref{sec:ladder}) and the per-claim
provenance map (\S\ref{sec:practice-provenance}) are the two practices
that bear the load. A citation cannot reach \textsc{lit-retrieved}
without a URL or DOI that resolves; cannot reach \textsc{ai-confirmed}
without a quoted snippet from the resolved source preserved in
\texttt{doc/sources.md}; cannot reach \textsc{lit-read} without a human
who has read it. Each rung is a state in the FSM, not a checkbox: an
elevation must be accompanied by an artefact in the repository.

\subsection{Claim--evidence drift}
\label{sec:fm-drift}

A single agent that writes prose, retrieves sources, and rates the
fit will, under any non-trivial workload, drift: the claim sharpens,
the evidence narrows, and the gap between the two becomes a hedge that
nobody notices. The handback discipline (\S\ref{sec:handback}) is the
mitigation: the writer cannot rate its own claims, the Source Analyzer
cannot rewrite the prose, and the Aligner can only file
hand-backs.

\subsection{Sloppification}
\label{sec:fm-sloppification}

Sloppification is the gradual loss of precision under iterative
LLM editing: technical terms get replaced by near-synonyms, named
constants get rounded, and quoted passages drift from their source. The
mitigation is the mirror discipline applied to quotations specifically
(\S\ref{sec:practice-mirror}): every quoted passage either appears
verbatim in a vendored source under \texttt{doc/sources/} or is held
as a literal in the provenance graph; a paraphrase that has crept into
quotation marks fails the audit pass.

\subsection{Model collapse}
\label{sec:fm-collapse}

Model collapse is the documented degradation of model output quality
when training corpora are increasingly populated by prior model output.
The mitigation in F(AI)\textsuperscript{2}R is indirect: by preserving
transcripts (\S\ref{sec:practice-transcripts}) and disclosing them
under \S\ref{sec:practice-disclosure}, downstream training corpora can
be filtered. The pattern does not prevent collapse; it does provide the
labelling that lets others avoid contaminating downstream pipelines.

\subsection{List-of-lists prose}
\label{sec:fm-listoflists}

A symptom of LLM-driven editing: prose that decomposes into nested
bullet lists, each list a paraphrase of the surrounding sentence,
because list output is a local optimum for many models. The
Readability \& Novelty Scrutinizer's defect registry has a category for
it; the Scientific Writer is responsible for the repair. The pattern
does not eliminate the symptom; it makes it visible.

\subsection{Hedging chains}
\label{sec:fm-hedging}

Hedging chains are sentences whose hedging clauses are so dense that
the contribution disappears: ``it may be argued, though some have
suggested otherwise, that under certain assumptions and pending
further work\dots''. The Readability scrutinizer flags hedge density
above a per-section threshold (default: one hedge per six to eight
sentences); the writer rewrites with the rule that each hedge must be
paired with a commitment in the same sentence. This rule is itself
unoriginal --- editors have enforced it for decades --- but it is
machine-checkable enough to delegate to an audit pass.

\subsection{Placeholder dishonesty}
\label{sec:fm-placeholder}

A polished surface that conceals which load-bearing parts remain
unverified is a worse failure than an obviously incomplete draft. The
mitigation is \emph{asset honesty over asset completeness}: a
placeholder is allowed and encouraged, but it must visibly identify
itself as AI-authored and pending. In \LaTeX{} the \texttt{todonotes}
package is used for inline placeholders; in figures, a labelled
``placeholder'' canvas is committed and replaced. A render that hides
its own gaps is a defect.

\subsection{Dual use}
\label{sec:fm-dualuse}

The same artefacts that make this work auditable also make it
adoptable by adversarial integrators. A pipeline of prompts, a
verification ladder, and a PROV-O schema are as much a recipe for
\emph{automating bad-faith scholarship at scale} as they are for
honest LLM-assisted writing. The pattern does not eliminate this
surface. It makes the surface visible in a form that can be reasoned
about, which is the most we are willing to claim.

\subsection{What the pattern does not address}
\label{sec:fm-residual}

Three failure modes are out of scope. \emph{Pre-publication leakage} ---
a model's training data containing the manuscript before it is meant
to --- is a vendor problem F(AI)\textsuperscript{2}R cannot reach.
\emph{Audience fatigue} --- readers who have stopped engaging with
disclosures because the boilerplate is identical at every venue --- is
a community problem the pattern can only signal toward.
\emph{Adversarial review} --- a reviewer who reads the prose and
ignores the graph, or vice versa --- is a sociological problem that no
artefact discipline can solve. We name these as residual.
