\section{Introduction}
\label{sec:intro}

The FAIR Guiding Principles for scientific data
management~\cite{wilkinson2016fair} were drafted for data, then
extended to research software (FAIR4RS~\cite{chuehong2022fair4rs}) and
to machine-learning models (FAIR4ML~\cite{ravi2024fair4ml}). Neither
extension yet covers the artefact classes that LLM-assisted scholarly
writing produces: exportable conversation transcripts, versioned
prompt files, tool-and-model version manifests, verification-status
ladders, structured redaction policies, and per-claim provenance maps.

This is not a hypothetical gap. Anyone now writing a paper with an
LLM in the room is producing those artefacts, whether or not they
preserve them. If they vanish, two things follow. First, the audit
trail is gone --- a reader cannot ask \emph{who wrote this sentence,
what evidence supports this claim, and how was the claim
verified?}~\cite{krebs2026obscurity}. Second, the work is more
exposed to the documented failure modes of LLM-assisted writing,
which sit on top of a reproducibility baseline that is already poor
in many fields~\cite{ioannidis2005,pineau2021reproducibility}\todo[inline]{verify}:
fabricated citations, drift between claim and evidence,
sloppification, and the gradual flattening of an argument as hedges
accrete. Reported base rates are not anecdotal: Walters and
Wilder~\cite{walters2023fabrication} report 55\% fabrication for
GPT-3.5 and 18\% for GPT-4 across general-purpose generation;
Magesh et al.~\cite{magesh2024legal} report 17--34\% hallucination
for retrieval-augmented legal-research tools. The numbers are not
directly comparable --- one grades full citations, the other grades
RAG-anchored answers --- but they are the same order of magnitude,
and it is the kind of number that turns a methodological
inconvenience into a disqualifying one.

The volume problem compounds the per-paper one. The fraction of
journal and arXiv submissions that contain LLM-generated or
LLM-assisted prose has risen sharply since
2023~\cite{liang2024mapping,vannoorden2023chatgpt}; detection studies
report measurable increases across computer science, biomedicine, and
the social sciences~\cite{kobak2024delving}. Editors at major venues
have publicly described AI-shaped text as a load they cannot review
out of submissions one paragraph at a time, and have asked for
provenance \emph{from} the authors instead of post-hoc detection
\emph{against} them~\cite{else2023chatgpt,thorp2023chatgpt}\todo[inline]{verify}. Paper mills, fabricated
literatures, and the boilerplate flattening of disclosure pages are
the symptoms; what is missing is a discipline that authors can adopt
on day one rather than a forensic regime that reviewers apply on the
back end. Detection alone is, on present evidence, an arms-race the
reviewer side does not
win~\cite{sadasivan2023reliably}\todo[inline]{verify}; an
authoring-side discipline is the cheaper, sturdier complement. The
pattern this paper proposes is meant to be that discipline.

\paragraph{The journal as distribution channel may be in decline.}
The strain is not only on the per-paper review pipeline; it is on
the format itself. The original concept of the journal --- a
periodical bundling vetted articles for distribution to subscribers
--- pre-dates the internet by three centuries, and the network has,
on aggregate, weakened the case for bundling. Preprint servers,
overlay journals, registered reports, open-science platforms, and
artefact-versioned repositories now cover much of what journals
historically did, often faster and more
openly~\cite{eisen2018preprints,tennant2016open}\todo[inline]{verify}.
The arrival of LLM-generated submissions on top of a system that
was already showing format strain is what tips a slow drift into a
visible crisis: editors face submission queues swollen by
AI-fabricated content~\cite{conroy2023sleuths}\todo[inline]{verify}
at the same time that authors face venue-policy
churn~\cite{aclrr_llm_policy,iclr_llm_policy,neurips_llm_policy}
and reviewers
face capacity exhaustion. We do not claim the journal is dead ---
the position would be over-strong, and many fields still need the
curation work journals do. We do claim that the assumption that the
journal is the unproblematic default for scholarly distribution can
no longer be taken for granted, and that an authoring-side
discipline whose primary outputs are the \emph{repository, the
provenance graph, and the manuscript bundled together} is at least
as well-suited to the emerging communication ecosystem as it is to
the legacy one. F(AI)\textsuperscript{2}R is designed to fit both.

We propose \emph{F(AI)\textsuperscript{2}R}, read \emph{F-A-I-A-I-R}:
the canonical FAIR axes with an additional \emph{(AI)} factor
multiplied through them, squared because each axis demands two passes
over every artefact, an authoring pass and an audit pass. An earlier
draft used \emph{FAIR4AI}; we discarded it after observing that
\emph{4AI heißt alles} --- a suffix denotes nothing about how the
principles change. Folding \emph{(AI)} into the acronym, multiplied
through, asserts the extension is internal to FAIR rather than
external to it.

\paragraph{Contribution.} The contribution is not the name. It is the
integration of eight individually unoriginal practices into a single
discipline: (1) transcript preservation; (2) verification-status
labelling; (3) per-claim provenance maps; (4) mirror discipline
between source and render; (5) recursive meta-process documentation;
(6) base-rate-anchored AI disclosure; (7) legal honesty about
authorship; and (8) FAIR alignment treated as a precondition, never
as a retrofit. None of the eight is novel on its own. Their joint
adoption, enforced by a ten-stage agent pipeline with strict
separation between prose-owning and audit-owning agents, is what we
claim is new. \emph{The pipeline is the team; the team is the
methodology.}

\paragraph{Roadmap.}
\S\ref{sec:position} states the F(AI)\textsuperscript{2}R position
in five numbered claims. \S\ref{sec:pattern} states the pattern
itself: the axes, the ten-stage pipeline, the verification ladder
as a finite-state machine, and the handback discipline.
\S\ref{sec:eight} treats the eight integrated practices.
\S\ref{sec:provenance-analysis} reports the audit pass on the live
graph; \S\ref{sec:evolution} narrates how the pipeline itself
evolved during cooperative writing. \S\ref{sec:failure-modes}
names the failure modes the pattern resists and is honest about
which it does not. \S\ref{sec:limits-and-objections} addresses the
dual-use surface, equity, and the standing objections;
\S\ref{sec:conclusion} states the conditions for
community-recommendation submission. Seven appendices supply the
verification-ladder specification, the prompt taxonomy, the
F(AI)\textsuperscript{2}R cell mapping, the process-evolution
chronology, the provenance-analysis queries, the implementation
notes, and the pattern's two forward-looking extensions (domain
ontologies and graph-as-research-infrastructure).
