\section{Limits, dual-use, and standing objections}
\label{sec:limits-and-objections}

The pattern is not free. We owe the reader an explicit account of
what F(AI)\textsuperscript{2}R consumes, what it depends on, and
what the standing objections to it look like.

\paragraph{Resource cost and frontier-model dependence.}
Every authoring pass and every audit pass burns compute, and the
verification ladder amplifies this: a contested claim is read by
an agent, by a human, sometimes by a second agent during
readability review, and again by the FAIR aligner. The energy and
water footprint of frontier-model inference is non-trivial and is
paid by someone, somewhere, regardless of who pays for the API
call~\cite{strubell2019energy,patterson2021carbon,luccioni2024power,li2023thirsty}\todo[inline]{verify}.
We add the following request to the sixth integrated practice
(\S\ref{sec:eight}, item~6): disclosure should
\emph{include the order-of-magnitude inference budget consumed by
the manuscript}, declared at submission time alongside the rung
distribution. Effective behaviour also depends on the underlying
model and shifts as models change~\cite{chen2023drift}\todo[inline]{verify};
the transcripts of practice~1 and the version manifest in
\texttt{doc/provenance.ttl} are partial mitigation, and any
\texttt{fair2r:Activity} record should carry an explicit
\texttt{fair2r:modelId} property pinned to a stable identifier so
that a deprecated agent at minimum has a known headstone.

\paragraph{Equity.}
Research groups without access to frontier models cannot apply
this pattern at the same effective rung; the verification ladder
requires an AI agent at \textsc{ai-confirmed} to actually
understand the source it has read, and the floor for ``understands''
is set by the state of the art. F(AI)\textsuperscript{2}R
\emph{describes} the work; it does not \emph{equalise} access to
it~\cite{bender2021stochastic,birhane2022values}\todo[inline]{verify}.
We name this honestly and offer two partial mitigations:
smaller-model fallbacks for routine audit passes (layout,
hedging-density, list-of-lists detection) that do not need
frontier-grade reasoning; and the structural property that an
honestly-declared lower-rung distribution at submission time is
preferable to a forged higher-rung distribution from a
richer group, with the verification ladder making the difference
legible to the reviewer.

\paragraph{Adjacent proposals and adjacent vocabularies.}
A concurrent maximalist proposal, the \emph{Agent-Native Research
Artifact} (Ara) of \textcite{liu2026ara}, replaces the narrative
paper with a four-layer package --- scientific logic, executable
code with full specifications, an exploration graph of failed
branches, and evidence grounding every claim --- and reports
reproduction-success and QA-accuracy gains on PaperBench and
RE-Bench. F(AI)\textsuperscript{2}R declines that bargain. We
keep the human-readable manuscript as the primary artefact and
ride a PROV-O graph alongside it; Ara's exploration layer is a
strict superset of our logbook plus claim-graph, and indicates a
credible upper bound for what the same substrate could carry if
a venue ever demanded it.

F(AI)\textsuperscript{2}R is also sometimes mistaken for a
dataset-metadata proposal. It is not. The dataset-metadata slot
is already occupied by MLCommons Croissant~1.0
\cite{benjelloun2024croissant}, a schema.org/JSON-LD vocabulary
for ML datasets and their responsible-AI annotations. Croissant
describes the \emph{thing learned from}; F(AI)\textsuperscript{2}R
describes the \emph{authoring process} that produces a paper.
The two compose: a Croissant record can be referenced from a
\texttt{fair2r:Source} triple, but neither schema subsumes the
other.

\paragraph{Standing objections.} A position paper that does not
name the objections it expects has not committed to a position.

\emph{``This is just FAIR with extra steps.''}
Two prior re-readings (FAIR4RS~\cite{chuehong2022fair4rs},
FAIR4ML~\cite{ravi2024fair4ml}) judged that adding sub-principles
for new artefact classes was preferable to absorbing them. We
follow that precedent; the audit pass on \textbf{Reusable} is the
load-bearing case --- replaying a writing pipeline is materially
different from replaying a build.

\emph{``Eight practices on day one is too expensive.''} Eight is
the \emph{minimum}: the failure-modes table
(Table~\ref{tab:failure-modes}) shows that dropping any one
exposes a specific failure mode the others do not cover. The
\textsc{ai-confirmed} rung was introduced exactly to attack the
verification-overhead bottleneck (\S\ref{sec:ladder}).

\emph{``Provenance graphs do not prevent fraud.''} Correct, and
we never claim otherwise. The graph aims to make fraud
\emph{detectable by inspection}: a bad-faith adoption leaves a
graph whose edges contradict the prose, whose claims reference
non-existent sources, or whose
\texttt{prov:wasGeneratedBy} activities have no transcript. Each
of those is a checkable defect; the contrast is with unstructured
AI-assisted writing, where the same fraud leaves no artefact a
third party can point at.

\emph{``You cannot check the model's reasoning, only its
output.''} Yes; we never claim otherwise. The discipline makes
the \emph{output's relationship to the input} checkable, even
when the model's internals are not. That is what the verification
ladder, the Source Analyzer hand-off, and the human-reader rung
are for.

\emph{``A position paper should not also be a worked example.''}
The fifth integrated practice (\S\ref{sec:eight}, item~5) requires
that a paper which prescribes a discipline have been written
under it. The recursive case study is not an aesthetic choice; it
is how the position can be checked.

\bigskip
\noindent The objections F(AI)\textsuperscript{2}R cannot answer
--- pre-publication leakage, audience fatigue, adversarial review
--- sit in \S\ref{sec:fm-residual} as residual.
