\section{The F(AI)\texorpdfstring{\textsuperscript{2}}{2}R Pattern}
\label{sec:pattern}

\subsection{Axes: F(AI)\texorpdfstring{\textsuperscript{2}}{2}R}
\label{sec:axes}

\input{figures/axes}

\input{figures/ai-squared-grid}

We retain the four canonical FAIR axes and read each in two passes
(Figure~\ref{fig:axes}). The (AI)\textsuperscript{2} factor itself
unpacks into the four cells of writer-vs-reader
(Figure~\ref{fig:ai-squared-grid}): a manuscript that will be both
written and read by humans only needs the manuscript proper, while a
manuscript whose readers include AI agents needs every artefact in
the F(AI)\textsuperscript{2}R bundle. The authoring pass (AI\textsubscript{1})
produces the artefact for that axis. The audit pass
(AI\textsubscript{2}) verifies the artefact against the axis after
the fact, never in the same role and never in the same commit. The
two passes are what justifies the squared notation. The (AI) factor
is multiplied through every axis.

\paragraph{F.}
Authoring mints persistent identifiers (DOI via Zenodo, ORCID for
authors), populates machine-readable metadata
(\texttt{CITATION.cff}, \texttt{codemeta.json},
\texttt{.zenodo.json}), and assigns IRIs to per-claim entities. Audit
verifies the metadata surfaces agree, no IRI is dangling, and every
manuscript-cited claim has a graph entry.

\paragraph{A.}
Authoring clones-and-builds: a third party with the public repository
and a documented toolchain reproduces the PDF. Audit verifies the
licence pair (MIT for code/RDF, CC-BY-4.0 for prose), that paywalled
sources have been vendored where licensable, and that the redaction
policy covers what it claims.

\paragraph{I.}
Authoring commits the provenance graph as PROV-O over Turtle in
\texttt{doc/provenance.ttl}. Audit validates the Turtle, runs SHACL
shapes if defined, and verifies every section has a
\texttt{prov:Activity} record that produced it.

\paragraph{R.}
Authoring commits the per-role prompts under \texttt{agents/}, the
build target under \texttt{paper/Makefile}, and the logbook under
\texttt{doc/logbook.md}. Audit replays the pipeline on a clean clone
and reports drift.

\subsection{The pipeline as the team}
\label{sec:pipeline}

\input{figures/pipeline}

Ten stages, coordinated by an Orchestrator, with strict handback
between prose-owning and audit-owning agents
(Figure~\ref{fig:pipeline}):

\begin{description}[leftmargin=1.2em,itemsep=2pt]
    \item[0. Orchestrator.] Inspects state, dispatches one downstream
        stage at a time. Never edits prose.
    \item[1. Research.] Per-section literature audit. Produces sources
        and transcripts under \texttt{doc/sources/} and
        \texttt{doc/transcripts/}.
    \item[1.5. Source Analyzer.] Promotes
        \textsc{lit-retrieved}\,$\to$\,\textsc{ai-confirmed}; never crosses into
        \textsc{lit-read}.
    \item[2. Scientific Writer.] Owns prose, register, \LaTeX. Does
        not change claims or introduce citations without a verified
        source.
    \item[3. Illustration.] Owns figures; halts on load-bearing values
        still at \textsc{lit-retrieved}.
    \item[4. Layout Scrutinizer.] Reads the rendered PDF, files a
        Layout Defect Registry under \texttt{doc/handbacks/}.
    \item[5. Readability \& Novelty Scrutinizer.] Reads the markdown,
        flags repetition, list-of-lists prose, hedging chains, and
        unsupported novelty. Files a registry.
    \item[6. Aligner (FAIR).] Audits traceability across paper,
        sources, README, and registries. Routes hand-backs; never
        edits content.
    \item[7. Provenance Curator.] Sole writer of
        \texttt{doc/provenance.ttl}; refuses to invent triples.
    \item[8. Site Agent.] Owns the public site; refuses to render
        without explicit publication consent. The site never
        introduces a new claim.
\end{description}

The agent that owns prose cannot mark its own claims as verified. The
agent that audits layout cannot edit prose. The agent that curates
provenance cannot write \LaTeX. \emph{Specialisation reduces
hallucination at the seam.}

\subsection{The verification ladder as an FSM}
\label{sec:ladder}

\input{figures/ladder-fsm}

Every claim carries a verification state. The states are organised
as a finite-state machine (Figure~\ref{fig:ladder-fsm}), monotone
non-decreasing modulo explicit retraction:

\begin{quote}\small
\textsc{unverified-external}
\(\to\) \textsc{needs-research}
\(\to\) \textsc{lit-retrieved}
\(\to\) \textsc{ai-confirmed}
\(\to\) \textsc{lit-read}
\end{quote}

\noindent \textsc{lit-retrieved}: a candidate URL or DOI that
resolves. \textsc{ai-confirmed}: an AI agent has read the source and
confirmed it supports the claim, with a quoted snippet preserved in
\texttt{doc/sources.md}. \textsc{lit-read}: a human has read the
source and confirmed both that it supports the claim and that the
inference from source to claim is sound. A claim may invoke a source
at its current rung or below; load-bearing or contested claims still
require \textsc{lit-read}.

The \textsc{ai-confirmed} rung exists because verification overhead
is real and dominates writer cycles. Most retrieved entries are
obvious from the abstract; a small minority anchor contested claims;
the single-tier gate forced the human reader through both populations
at the same cost. Splitting the work between the Source Analyzer
(cheap) and the human (expensive) is the lever that makes the
practice economically tenable.

\paragraph{Crosswalk.} Retrieval depth maps to PRISMA 2020:
\textsc{lit-retrieved}\,=\,Identification, \textsc{ai-confirmed}\,=\,Screening,
\textsc{lit-read}\,=\,Included. Invocation strength maps to GRADE: a
claim invoked at \textsc{lit-read} carries a GRADE certainty floor of
\textsc{moderate}; a claim at \textsc{ai-confirmed} can carry at most
\textsc{low}. The mapping is a floor, not a ceiling.

\subsection{Handback and mirror disciplines}
\label{sec:handback}

The four scrutinizing stages --- Source Analyzer, Layout Scrutinizer,
Readability \& Novelty Scrutinizer, Aligner --- never edit source.
They write \emph{Defect Registries} under
\texttt{doc/handbacks/}; the repair lands in a subsequent commit by
a prose-owning agent. The cost is one extra commit boundary per
cycle. The payoff is that ``what did the scrutinizer actually flag?''
has a single canonical answer in the git history.

Every primary artefact has two mirrored representations: the source
under version control, and the rendered output. The pair is enforced
for the manuscript, the provenance graph, the prompts, and the
logbook. Drift between source and render is detectable; the build
either succeeds or fails.

The pattern, in one sentence: \emph{the manuscript, the provenance
graph, and the logbook are primary artefacts and must remain
consistent and up to date at all times}. A commit that updates one
and leaves the others lagging is a defect, even if the build passes.

\subsection{Extensibility through domain ontologies}
\label{sec:domain-ontologies}

PROV-O is the spine; it does not have to be the only ontology in
the graph. The schema we have introduced is compositional: any
\texttt{fair2r:Claim} can carry additional triples that ground its
\emph{values} in a domain ontology, alongside the
\texttt{prov:wasDerivedFrom} edge that grounds its
\emph{evidence}. The two axes are orthogonal --- provenance asks
\emph{where did this claim come from?}, domain ontologies ask
\emph{what does the value of this claim mean?} --- and a manuscript
that answers both is more reproducible than one that answers only
the first.

Two examples concrete enough to plug in today:

\begin{description}[leftmargin=1.2em,itemsep=2pt,topsep=4pt]
    \item[\textbf{Sensor data: SOSA / SSN.}] The W3C Semantic Sensor
        Network ontology and its lightweight core
        SOSA~\cite{janowicz2019sosa}\todo[inline]{verify} expose
        \texttt{sosa:Observation}, \texttt{sosa:Sensor},
        \texttt{sosa:FeatureOfInterest}, and the relations between
        them. A \texttt{fair2r:Claim} that asserts a measurement can
        carry a \texttt{prov:wasDerivedFrom} pointing at the
        underlying \texttt{sosa:Observation}, and the audit pass
        gains the ability to ask \emph{is this measurement well-
        formed?} as a SHACL query rather than a paragraph of
        prose.
    \item[\textbf{Units of measure: OM-2 / QUDT.}] The OM-2 ontology
        of units of measure~\cite{rijgersberg2013om}\todo[inline]{verify}
        and the QUDT vocabulary~\cite{qudt}\todo[inline]{verify}
        expose \texttt{om:Quantity}, \texttt{om:Unit},
        \texttt{qudt:hasUnit}, and the dimensional relations between
        units. A claim that asserts a numeric value can carry the
        unit as an explicit triple rather than as a typographic
        decoration. Unit consistency is then a graph query, not a
        manuscript-reading exercise.
\end{description}

\noindent The class of failure modes this addresses is the one
prose alone cannot reach: a value that is internally coherent in
its prose and its citation but \emph{wrong in its units}, or
\emph{wrong in the dimensional sense it implies}. A reviewer
catching such a defect today reads the figure caption and the
methods section and triangulates; an audit pass with a domain
ontology reads the graph and the SHACL shapes and reports.

The recommendation is open. F(AI)\textsuperscript{2}R does not
prescribe specific domain ontologies because the right choice is
discipline-specific. It prescribes that, where a discipline already
has a \emph{konkrete standardisierte Methode} expressed as an
ontology, the F(AI)\textsuperscript{2}R repository should adopt it
and link from \texttt{fair2r:Claim} entities to its terms. NFDI4Ing
and the Helmholtz Metadata Collaboration are the nearest two
communities to this paper that maintain such ontologies for
engineering and metadata respectively, and the integration cost is
low. The ceiling is the discipline's; the floor is what
F(AI)\textsuperscript{2}R asks for.

\subsection{Looking forward: provenance graphs as research infrastructure}
\label{sec:research-infrastructure}

The argument of \S\ref{sec:domain-ontologies} generalises. The
discipline that pioneered semantic-tagging-as-norm under
data-volume pressure is bioinformatics: faced with what was, at the
time, an obscene amount of sequence data, the field standardised
identifier and annotation schemas (the Gene
Ontology~\cite{ashburner2000go}\todo[inline]{verify} being the most
visible example) so that genes, proteins, and their relations could
be referenced unambiguously across labs, papers, and databases. The
ontology was not a research output; it was a precondition for
research that scaled. The community's investment in semantic
tagging was, in retrospect, what made the next decade of
bioinformatics possible.

We claim a closer analogy than is sometimes recognised. LLM-assisted
scholarly writing is producing artefacts at a volume that the
unaided reviewer pipeline cannot absorb (\S\ref{sec:intro}). The
response will, on present trajectory, look more like the
bioinformatics response than like the editorial response: not new
detection tools but new shared schemas, ontologies, and
graph-interconnected metadata that let claims, sources, and
activities reference each other across institutional boundaries.
\textbf{Future research success may come to depend less on the
solitary craft of a single manuscript and more on three things the
F(AI)\textsuperscript{2}R schema is designed to support:} (i)
provenance models that turn published claims into machine-readable
hypotheses for the next study; (ii) graph interconnectivity across
institutional knowledge graphs --- the Helmholtz cross-centre
graph~\cite{curdt2025hmc} is the proximate example, and a
fully-instrumented version would let a claim in one centre's
manuscript be linked, queried, and challenged by claims in another's
without leaving the linked-data layer; (iii) clever experimental
validation that uses the graph to identify the next testable
deviation from the published consensus.

That trajectory is a position the paper is willing to take and a
direction the eight integrated practices serve. The verification
ladder feeds (i); the per-claim provenance map and the
domain-ontology side-arms feed (ii); the recursive case study and
the audit pass feed (iii). None of this requires F(AI)\textsuperscript{2}R
to be the standard. It requires only that some standard with these
properties be adopted, and the present paper is one offer among
several the community can refine.
