\section{Position}
\label{sec:position}

This paper takes a position. The five claims below are the position
in compressed form. They are stated here, defended in
\S\ref{sec:background}--\S\ref{sec:eight}, exercised on the present
manuscript in \S\ref{sec:impl}--\S\ref{sec:evolution}, and tested
against the failure modes a thoughtful reader will name in
\S\ref{sec:objections}.

\begin{enumerate}[leftmargin=1.6em,itemsep=2pt,topsep=4pt]
  \item \textbf{FAIR is insufficient for LLM-assisted scholarly
        writing in its current form.} The fifteen sub-principles of
        Wilkinson and colleagues~\cite{wilkinson2016fair} were drafted
        for data, and the FAIR4RS~\cite{chuehong2022fair4rs} and
        FAIR4ML~\cite{ravi2024fair4ml} extensions were drafted for
        software and models. None covers the artefact classes that
        LLM-assisted writing produces: transcripts, prompts, version
        manifests, verification ladders, redaction policies, per-claim
        provenance maps. A third extension is required.
  \item \textbf{The required extension is multiplicative, not
        appended.} Folding \emph{(AI)} into the FAIR acronym ---
        \emph{F(AI)\textsuperscript{2}R}, with the (AI) factor passed
        through every axis and squared because each axis demands two
        passes (authoring and audit) --- asserts that the change is
        \emph{internal} to FAIR rather than external to it. A
        \emph{4AI} suffix denotes nothing about how the principles
        change; \emph{4AI heißt alles}.
  \item \textbf{Eight integrated practices are the minimum
        discipline.} Transcript preservation, verification-status
        labelling, per-claim provenance maps, mirror discipline,
        recursive meta-process documentation, base-rate-anchored
        disclosure, legal honesty about authorship, and FAIR
        alignment as precondition. None of the eight is novel on its
        own. The integration is the contribution, and adopting any
        subset is, on the failure modes the literature has
        documented, insufficient.
  \item \textbf{The pipeline is the team; the team is the
        methodology.} A ten-stage agent pipeline with strict handback
        between prose-owning and audit-owning agents
        (\S\ref{sec:pipeline}) is the cheapest known mechanism that
        makes the eight practices enforceable rather than aspirational.
        Specialisation reduces hallucination at the seam.
  \item \textbf{Naming the practice is the first step toward
        contesting it.} F(AI)\textsuperscript{2}R is offered as a
        position to be argued with --- contestable in the open
        literature, replayable on a clean clone of the public
        repository, and revisable as the capability of the underlying
        models, the disclosure norms of the venues, and the licensing
        posture of the corpora that train the next models all move
        under us.
\end{enumerate}

\noindent The remainder of the paper defends each of the five
claims. A reader who disagrees with any one of them is invited to
say so and to bring evidence; the artefact discipline this paper
prescribes is engineered specifically so that the disagreement can
be checked against the same graph the position rests on.
