\section{Statement of authorship and AI use}
\label{sec:authorship-statement}

This statement adapts the \emph{Statement of independence} from the
methodological ancestor~\cite{krebs2026obscurity} for a paper that
is, by contrast, institutionally embedded.

\paragraph{Author.}
Florian Krebs (ORCID
\href{https://orcid.org/0000-0001-6033-801X}{0000-0001-6033-801X})
is the sole human author. The work is conducted at the German
Aerospace Center (DLR / \emph{Deutsches Zentrum für Luft- und
Raumfahrt}), Zentrum für Leichtbauproduktionstechnologie (ZLP),
Augsburg. The author is responsible for every claim in the
manuscript and for the integrity of the artefacts the manuscript
ships with.

\paragraph{Institutional context.}
DLR is a member of the Helmholtz-Gemeinschaft Deutscher
Forschungszentren. Two Helmholtz initiatives are relevant to this
work and are acknowledged here in addition to the author's
affiliation: the Helmholtz Metadata Collaboration (HMC), whose
mandate is FAIR metadata across the Helmholtz association, and
NFDI4Ing, the DFG-funded research-data infrastructure for
engineering sciences. F(AI)\textsuperscript{2}R intersects both
mandates and is offered to both communities for refinement.

\paragraph{AI use.}
The manuscript was drafted in cooperation with a Large Language
Model (Anthropic Claude, model identifier
\texttt{claude-opus-4-7}) operating under the per-role prompts
committed to \texttt{agents/} and the harness rules in
\texttt{CLAUDE.md}. The model is acknowledged here per the seventh
integrated practice (\S\ref{sec:eight}, item~7); it is \emph{not}
an author. The legal position is consistent across the
jurisdictions surveyed: in Germany
(\S2 \emph{UrhG}~\cite{urhg2}), in the United States
(USCO 2023~\cite{usco2023ai}), and at the major medical journals
via the \textsc{ICMJE}~\cite{icmje2023} and at \emph{Science} via
the editor-in-chief~\cite{thorp2023chatgpt}, AI cannot hold
copyright, cannot consent to publication, and cannot be held
accountable. The provenance graph in \texttt{doc/provenance.ttl}
records, for each section, the activity that produced it, the
agents (human and AI) that participated, and the verification rung
of every named claim. The \texttt{CITATION.cff} lists humans only.

\paragraph{Data and code availability.}
The repository\footnote{\url{https://github.com/noheton/f-ai-r}}
contains the \LaTeX{} sources, the bibliography, the agent prompts,
the build pipeline, the static-site generator, the PROV-O graph,
the logbook, and the redaction and consent policies. Code, prompts,
and RDF are released under MIT; manuscript prose is released under
CC-BY-4.0. The DLR design-system assets vendored under
\texttt{site/static/dlr/} are subject to upstream licensing terms.

\paragraph{Conflict of interest.}
None declared. The author has no financial relationship with any
LLM vendor.

\paragraph{Funding.}
This work was carried out in the course of the author's regular
research activity at DLR. No targeted external funding supported
the writing of this paper.

\paragraph{Acknowledgements.}
We thank the maintainers of the underlying open-source toolchain
--- TeX Live, \texttt{latexmk}, \texttt{biber}, \texttt{rdflib},
the Python ecosystem, GitHub Actions --- without which the pattern
would be more expensive than it is. The pattern here distilled was
developed on the source repository
\texttt{noheton/Obscurity-Is-Dead}~\cite{krebs2026obscurity} by the
same author; the present paper is the abstraction that strips the
source's domain content and publishes the methodological layer.
