\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage[backend=biber,style=numeric,sorting=none]{biblatex}
\usepackage{todonotes}

\addbibresource{references.bib}

\title{F(AI\textsuperscript{2})R: FAIR Research with AI in the Loop, Twice}
\author{TBD\thanks{TBD}}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
The FAIR principles (Findable, Accessible, Interoperable, Reusable) were
formulated in an era when scholarly artefacts were authored by humans and
consumed, eventually, by other humans. Generative AI has changed both ends
of that pipe: Large Language Models now help author papers and will, soon,
be the dominant first-pass readers of them. We argue that FAIR therefore
needs a second copy of AI inside the loop --- once as \emph{assistant} in
authoring, once as \emph{auditor} of provenance --- and we sketch a
working pattern, F(AI\textsuperscript{2})R, in which both passes are
captured in a PROV-O graph that ships with the manuscript. We demonstrate
the pattern on this very paper: every section, figure, and citation is
linked, by Turtle triple, to the agent (human or model) that produced it
and to the source it draws on. We discuss what this buys (replayability,
contestability, machine-grade attribution) and what it costs
(prompt drift, hallucinated citations, the temptation to launder uncertainty
through provenance metadata).
\end{abstract}

\section{Introduction}
\label{sec:intro}
\todo[inline]{Motivate: FAIR is fifteen years old; LLMs broke the
authorship/readership symmetry; we need a refresh, not a replacement.}

\section{Background: FAIR and PROV-O}
\label{sec:background}
\todo[inline]{Recap FAIR (Wilkinson et al., 2016) and PROV-O. Position
F(AI\textsuperscript{2})R as additive, not revisionist.}

\section{The F(AI\textsuperscript{2})R Pattern}
\label{sec:pattern}
\todo[inline]{Two AI passes: assistant and auditor. Roles, prompts,
artefacts. Show diagram.}

\section{Implementation in This Repository}
\label{sec:impl}
\todo[inline]{Describe directory layout, agent set, Makefile, CI, the
provenance graph, the logbook.}

\section{Discussion}
\label{sec:discussion}
\todo[inline]{Limitations: prompt drift, hallucinated citations, the
``provenance theatre'' failure mode.}

\section{Related Work}
\label{sec:related}
\todo[inline]{Provenance-aware writing tools, FAIR-for-software, model
cards, datasheets, the recent literature on AI co-authorship.}

\section{Conclusion}
\label{sec:conclusion}
\todo[inline]{Summarise; release this repository as the worked example.}

\printbibliography

\end{document}
